\chapter{Solvers}

\dolfin{} provides a number of pre-defined PDE solvers (called
``modules'' in the source structure) by default. The solver interface
is intentionally very simple to facilitate users writing their own
solvers. These are the pre-defined solvers:

\begin{enumerate}
\item
Poisson
\item
Convection-Diffusion
\item
Navies-Stokes
\item
Elasticity
\end{enumerate}

A solver for a PDE should provide the following interface:

\begin{enumerate}
\item
a constructor which takes a mesh, equation coefficients and possibly
additional data.
\item
a \texttt{solve()} method which solves the equation given the
specified data.
\item
a static \texttt{solve()} function which constructs and solves the
equation.
\end{enumerate}


\fixme{List solvers, then present in detail, include lots of nice images with
solver output}

%------------------------------------------------------------------------------
\section{Poisson's equation}
\index{Poisson's equation}

Poisson's equation with Dirichlet and homogenous Neumann boundary
conditions:

\begin{equation} \label{eq:poisson}
  \begin{array}{rcl}
    - \Delta u &=& f \quad \mbox{in } \Omega, \\
    u &=& g_D \quad \mbox{on } \Gamma_1, \\
    - \partial_n u &=& 0 \quad \mbox{on } \Gamma_2
  \end{array}
\end{equation}

The variational formulation is given by

\begin{equation} \label{eq:poisson-varform}
\begin{array}{rcl}
  \int_{\Omega} \nabla u \cdot \nabla v \, dx &=&
  \int_{\Omega} fv \, dx 
  \quad \forall v.
\end{array}
\end{equation}

The boundary conditions are enforced strongly and thus don't appear in
the variational formulation.

\subsection{Usage}

The API for the Poisson solver:

\begin{verbatim}
// Create Poisson solver
PoissonSolver(Mesh& mesh, Function& f, BoundaryCondition& bc);
    
// Solve Poisson's equation
void solve();

// Solve Poisson's equation (static version)
static void solve(Mesh& mesh, Function& f, BoundaryCondition& bc);
\end{verbatim}

A simple example of using the solver:

\begin{verbatim}
int main()
{
  Mesh mesh("mesh.xml.gz");
  MyFunction f;
  MyBC bc;
  
  PoissonSolver::solve(mesh, f, bc);
  
  return 0;
}
\end{verbatim}

Where ``f'' is a Function specifying the right-hand side of the
equation and ``bc'' is a BoundaryCondition.

\subsection{Performance}

The solver is an illustrative example and performance has not been a
goal. It uses a GMRES linear solver, where a multi-grid linear solver
would give optimal performance.

\subsection{Limitations}

The solver is meant to be the simplest example solver, and therefore
some simplifications have been made. Typically the general form of
Poisson's equation includes a diffusion coefficient which has been
omitted here.

%------------------------------------------------------------------------------
\section{Convection--diffusion}
\index{convection--diffusion}

The convection-diffusion equation with Dirichlet and homogenous
Neumann boundary conditions is given by:

\begin{equation} \label{eq:convdiff}
  \begin{array}{rcl}
    \dot{u} + b \cdot \nabla u - \nabla \cdot (a \nabla u) &=& f \quad \mbox{in } \Omega \times (0,T], \\
    u &=& g_D \quad \mbox{on } \Gamma_1 \times (0,T], \\
    - \partial_n u &=& 0 \quad \mbox{on } \Gamma_2 \times (0,T], \\
    u(\cdot,0) &=& u_0 \quad \mbox{in } \Omega,
  \end{array}
\end{equation}

where the convection is given by the vector $b = b(x,t)$ and the
diffusion is given by $a = a(x,t)$.

The variational formulation is:

\fixme{Stabilized convection-diffusion}

This is a stabilized FEM-formulation, so the solver can handle
convection-dominated problems.

The time integration is done using cG(1) (Crank-Nicolson).

\subsection{Usage}

The API for the convection-diffusion solver:

\begin{verbatim}
// Create convection-diffusion solver
ConvectionDiffusionSolver(Mesh& mesh, Function& w, Function& f,
  BoundaryCondition& bc);
    
// Solve convection-diffusion
void solve();
    
// Solve convection-diffusion (static version)
static void solve(Mesh& mesh, Function& w, Function& f,
  BoundaryCondition& bc);
\end{verbatim}

A simple example of using the solver:

\begin{verbatim}
int main()
{
  dolfin_output("curses");

  Mesh mesh("dolfin.xml.gz");
  Convection w;
  Source f;
  MyBC bc;

  ConvectionDiffusionSolver::solve(mesh, w, f, bc);
  
  return 0;
}
\end{verbatim}


\subsection{Performance}

There are no particular performance issues with the solver. GMRES is
used for solving the linear system.

\subsection{Limitations}

Currently many coefficients (such as diffusitivity) are not
user-definable, they need to be exposed by the interface.

%------------------------------------------------------------------------------
\section{Incompressible Navier--Stokes}
\index{Navier--Stokes}
\index{incompressible Navier--Stokes}

Write introduction here, equations etc.

\subsection{Usage}

Present API of solver and give an example.

\subsection{Performance}

Write something about the performance of the solver.

\subsection{Limitations}

Write something about the limitations of the solver.

%------------------------------------------------------------------------------
\section{Elasticity}

Navier's equations of elasticity with Dirichlet and homogenous Neumann
boundary conditions:

\begin{equation*}
\label{classicalelast}
\begin{split}
u &= x - X,\\
\dot{u}-v &= 0\quad\mbox{in } \Omega^0,\\
\dot{v} - \nabla \cdot \sigma& =f \quad\mbox{in } \Omega^0,\\
\sigma &= E\epsilon (u) = E(\nabla u^\top +\nabla u)\\
E\epsilon &= \lambda tr(\epsilon) I + 2\mu \epsilon,\\
v(0,\cdot ) &= v^0,\quad u(0,\cdot ) = u^0\quad\mbox{in } \Omega^0, \\
u &= g_D \quad \mbox{on } \Gamma_1 \times (0,T], \\
- \partial_n u &= 0 \quad \mbox{on } \Gamma_2 \times (0,T]
\end{split}
\end{equation*}

The variational form:

\begin{equation} \label{eq:elasticity-varform}
\begin{array}{rcl}
  \int_{\Omega} \dot{v} w \, dx = \int_{\Omega} -\sigma(u) \epsilon(v) + f w \, dx,
  \quad \forall w.
\end{array}
\end{equation}

The time integration is done using dG(0) (backward Euler).

The mass matrix appearing from $\int_{\Omega} \dot{v} w dx$ is lumped
(equivalent to computing it using nodal quadrature).

\subsection{Usage}

Present API of solver and give an example.

\subsection{Performance}

Write something about the performance of the solver.

\subsection{Limitations}

Write something about the limitations of the solver.
